---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
  library("data.table");
  library("xtable");
  activity <- data.table(read.csv("data/activity.csv"));
```


## What is mean total number of steps taken per day?

- Total number of stpes per day
```{r num_steps, results="asis"}
num_of_steps = xtable(activity[!is.na(steps),list( num_of_steps = sum(steps)), by=date]);
print(num_of_steps, type="html", include.rownames = FALSE);
```

- Total number of steps per day (histogram)
```{r hist_num_steps,  fig.height=4, fig.width=8, fig.align='center'}
hist(num_of_steps$num_of_steps, main = "Number of steps per day", xlab = "number of steps");
```

- Mean and Median for  number of stpes per day 
```{r mean_median_steps, results='asis'}
num_of_steps = xtable(activity[!is.na(steps),list(mean_num_steps = mean(steps),  median_num_steps = median(steps)), by=date]);
print(num_of_steps, type="html", include.rownames = FALSE);
```

## What is the average daily activity pattern?

- Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken

```{r time_series_plot}
  res <- activity[,mean(steps, na.rm = TRUE), by=interval];
  plot(res$interval, res$V1, xlab = "Interval", ylab = "Number of steps", main = "Time series plot", type = "l");

```

- Maximum number of steps interval

```{r max_num_steps_interval}
  res <- activity[,mean(steps, na.rm = TRUE), by=interval];
  paste("Maximam number of steps interval", res[res$V1 == max(res$V1)]$interval);
```

## Imputing missing values

- Total number of missing values in the dataset
```{r total_number_missing_steps}
  res <- activity[is.na(steps)];
  paste("Total Number missing", nrow(res));
```

- filling in all of the missing values in the dataset and creating dataset with field value (2 && 3)

```{r filling_missing}
mean_by_interval <- activity[,mean(steps, na.rm = TRUE), by=interval];
fill_activity <- activity;
fill_activity[is.na(steps)]$steps <- as.integer(mean_by_interval[interval == fill_activity[is.na(steps)]$interval]$V1)
```

- Total number of steps per day (histogram) (Imputing missing value )

```{r hist_num_steps_filled,  fig.height=4, fig.width=8, fig.align='center'}
hist(fill_activity[,sum(steps),by=date]$V1, main = "Number of steps per day (Imputing missing value)", xlab = "number of steps");
```

- Mean and Median for number of stpes per day (Imputing missing value )
```{r mean_median_steps_imputing, results='asis'}
num_of_steps = xtable(fill_activity[!is.na(steps),list(mean_num_steps = mean(steps),  median_num_steps = median(steps)), by=date]);
print(num_of_steps, type="html", include.rownames = FALSE);
```

Yes the value differ from the first part of assignment, median start to show up for some values instead of 0's.

## Are there differences in activity patterns between weekdays and weekends?

- New factor variable in the dataset with two levels – “weekday” and “weekend”

```{r, message=F, warning=F}
library("lubridate");
library("dplyr");
d <- ymd(activity$date);
wk <- weekdays(d);
is_week_end <- lapply(wk, function(x) { f <- "weekday"; if(x == "Sunday" || x == "Saturday") f <-"weekend"; f; });
activity <- cbind(activity, as.factor(unlist(is_week_end)));
activity <- rename(activity, type = V2);

```

- Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken

```{r time_series_plot_day_type, fig.width=8, message=F, warning=F}
  library(lattice);
  res <- activity[,mean(steps, na.rm = TRUE), by=c("type", "interval")];
  xyplot(res$V1 ~ res$interval   | res$type,  xlab = "Interval", ylab = "Number of steps", main = "Time series plot based on type", type = "l" );

```

